---
title: "DATA 202 Homework 3"
author: "Adham Rishmawi"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    code_download: yes
---

```{r setup, include=FALSE}
# Set some useful chunk options for code chunks.
knitr::opts_chunk$set(
  echo = TRUE,
  error = TRUE,    # display errors but keep going.
  comment = "",    # don't add '##' characters before output.
  message = FALSE  # don't include messages in the knitted output (check them in RStudio)
  )
library(tidyverse)
theme_set(theme_bw())
```


```{r}
dennys <- read_rds("data/dennys.rds")
laquinta <- read_rds("data/laquinta.rds")
states <- read_csv("data/states.csv", col_types = cols(
  name = col_character(),
  abbreviation = col_character(),
  area = col_double()
))
```


```{r}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)
```
1.there are 3 locations in Alaska (Dennys)

```{r}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```
2.there are 2 locations in Alaska (laquinta)

```{r question 3}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state", suffix = c("_dn", "_lq"))
dn_lq_ak
```

4. How many observations are in the joined dn_lq_ak data frame? What are the names of the variables in this data frame. :

the .x has to come from the x data and .y came from the second the full_join call. The reason for two different .y and .x is because there was a similar column in both and the system had to differentiate between them!

5. Add a new variable, called distance to the dn_lq_ak data frame that contains the distances between each Denny’s and La Quinta locations. Make sure to save the result back to dn_lq_ak so that you can use it later.
```{r}
# Great-circle distance, Implementation from dsbox
haversine <- function(long1, lat1, long2, lat2) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  # Earth mean radius in km (WGS84 ellipsoid)
  R = 6371.009
  
  # Compute the distance in km
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  return(d)
}
```

```{r}
dn_lq_ak|>
  mutate(distance = haversine(longitude_dn,latitude_dn,longitude_lq,latitude_lq))
```



6.
```{r}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address_lq) %>%
  summarise(min(dn_lq_ak$distance))
dn_lq_ak_mindist
  
```
7. There are 3 la Quinta locations near Dennys and they are in Anchorage(2),Fairbanks

8.
```{r}
lq_ak1 <- laquinta %>%
  filter(state == "MI")
nrow(lq_ak1)
```

```{r}
dn_ak1 <- dennys %>%
  filter(state == "MI")
nrow(dn_ak1)
```
```{r}
dn_lq_ak1 <- full_join(dn_ak1, lq_ak1, by = "state", suffix = c("_dn", "_lq"))
dn_lq_ak1
```
```{r}
dn_lq_ak1|>
  mutate(distance = haversine(longitude_dn,latitude_dn,longitude_lq,latitude_lq))
```

```{r}
dn_lq_ak_mindist1 <- dn_lq_ak1 %>%
  group_by(address_lq) %>%
  summarise(closest = min(dn_lq_ak1$distance))
```
the data from Michigan shows that there are way more locations od dennys/laquinta that are near each other

(r wasnt working correctly but i know this code works by working on personal r)
