---
title: "Data 202 -- HW 01"
author: "Adham Rishmawi"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    fig_height: 2.2
    fig_width: 4
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---
Spring 2022
```{r, setup, include = FALSE}
# load packages that are going to be used
require(tidyverse)   # this loads mosaic, ggformula, etc. too
require(ggformula)
require(mosaic)

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw(base_size=12))     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  echo = TRUE,      # for homework, always show R code (this is the default)
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small",   # slightly smaller font for code
  message = FALSE, warning = FALSE) # don't print warnings or messages in compiled document. So you MUST check them in RStudio!
```


##### the purpose of this program is to detect the optimal amount of rental bikes in each area for maxium effiecency and avoid delays for the customer.


###### This is data in relevance to the issue at hand (2011-2012)


```{r}
bikedata <- read_csv('~/data202/homework01/bikeshare-day-simple.csv', show_col_types = FALSE)
bikedata
```
this data set has 731 observations/occurrences and has 4 different variables at hand:
$date

$day_of_week

$workingday

$total_rides

```{r}
bikedata %>%
  ggplot() +
  aes(x = date, y = total_rides) +
  geom_point() +
  geom_smooth() +
  labs(
    x = "date", 
    y = "number of rides"
  )
```



```{r}
bikedata %>%
  ggplot() +
  aes(x = date, y = total_rides, col = workingday) +
  geom_point() +
  geom_smooth() +
  labs(
    x = "date", 
    y = "number of rides"
  )
```




interpretation: it seems like captialbike share's bikes are in more demand throughout the weekdays meaning that their products should be more in stock in the weekdays than the weekends!
  