---
title: "Stat 245 -- HW 1 "
author: "Adham Rishmawi"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 4
    fig_width: 9
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(tidyverse)   # this loads mosaic, ggformula, etc. too
require(ggformula)
require(mosaic)

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw(base_size=12))     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  echo = TRUE,      # for homework, always show R code (this is the default)
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small",   # slightly smaller font for code
  message = FALSE, warning = FALSE) # don't print warnings or messages in compiled document. So you MUST check them in RStudio!
```


```{r}
wrdata <- read_csv('https://sldr.netlify.app/data/sustainable-livelihoods.csv', show_col_types = FALSE)
```
##### How does gender densities differ in 3rd world countries?
```{r}
gf_dens( ~Gender, 
            color = ~Country,
           
            data = wrdata) 
```
I referred to the content by:

-looking at how to graph a density chart 

-understanding when to use gf_density and gf_dens

The key take away from my chart is to show the potential of a density chart yet state how i did it wrong!
My graph shows that the density of men triumphs the density of women in any of the 3rd world countries potential elaborating an issue of unequal gender division

criticism:  I believe my graph is missing a factor of time because it seems to try to a show a increase of density over a period of time but it might be also a wrong type of graph selection.

I did try to add some variables however it kept on giving me errors so this is my best attempt!


  