---
title: "Lab 3.2 - Visualization Composition"
author: "Adham Rishmawi"
date: "Spring Semester"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(scales)
library(gapminder)
```

This document presents some modifications on the health-and-wealth plot.

## Rebuilding the Health-and-Wealth Plot
Type or copy this code into the following code chunk one line at a time, running the code for each new line to see what it does. We&rsquo;ve started the process for you.


Here is the code for the plot shown in class on Wednesday:
```{r}
gapminder %>% 
  filter(year == 2007) %>% 
  ggplot() +
  aes(x = gdpPercap, y = lifeExp) +
  geom_point(alpha = .8) + 
  aes(color = continent) +
  aes(size = pop) + 
  scale_x_continuous(
    breaks = c(400, 4000, 40000),
    trans = "log10") +
  labs(x = "GDP per Capita") +
  labs(y = "Life Expectancy (years)") +
  labs(color = "Continent") +
  labs(size = "Population") +
  scale_size_area(labels = label_comma()) +
  theme_bw() + 
  annotation_logticks(sides = "b")
```



Notes:

- You can run the code chuck and see the output in the RStudio editor pane by pressing the green arrow (on the upper right of the chunk).

- Write a comment about what that line does using the `#` comment notation. The first 1.5 comments are written for you (e.g., `# Include only the records where...`).
  
- Omit the ` %>% ` (the *pipe* operator) or `+` at the end of a line until you're ready to enter the next line (e.g., you'll need to add `%>%` to the end of the filter line to go on).

- Practice using tab completion and the keyboard shortcut `Ctrl-Shift-M` to enter the ` %>% `.
    
- Look up the Help for a few of the functions by clicking on the function name and then pressing the `F1` key.
  
- Run the chunk. Observe what changed from what was there before.

## Modifying the Health-and-Wealth Plot

Here is a modified version of the previous plot.

🚧
```{r more-health-and-wealth}
gapminder %>% 
  #this is calling our main dataset and utlizing the pipe operator
  filter(year == 2002) %>% 
  #this filters years and onlyuses 2002 year datasheet
  ggplot() +
  aes(x = gdpPercap, y = lifeExp, color = continent, size = pop) +
  #this gives the aes specifications all in one line
  geom_point(alpha = .8) +
  scale_x_continuous(
    breaks = c(400, 4000, 40000),
    trans = "log10",
    labels = label_comma()) +
  #this creates the framework for the graph and refines the lables
  labs(x = "GDP per Capita",y = "Life Expectancy (years)", color ="Continent", size = "Population") +
  #this gives the strings for the labels and gives color and volume to country sizes
  scale_size_area(labels = label_comma()) +
  #separates labels with commas
  theme_dark() +
  #changes the theme of the graph to a darker theme
  annotation_logticks(sides = "b") +
  coord_cartesian(ylim = c(40,120))
  #changes the spread of data across the graph
```
<!-- This graph is a better display particularity because of a better theme that makes it visually more appropriate-->



Duplicate your code chunk here, change its name (e.g., `{r more-health-and-wealth}`), and try these extensions:

 

- Change the year to your birth year or another interesting year in history. Note: here are the available years:

 

```{r}

unique(gapminder$year)

```

 

- Try combining the `aes` lines together into a single `aes` command (with commas between the parts). Then try the same for the `labs`.

 

- Use commas in the `x` axis labels. (Find the part that adds commas to the `scale` for `area`. Use the same `labels` code in the part that already customizes other aspects of the `scale` for `x`.)

 

- Modify the scaling range of the y axis by adding this line `coord_cartesian(ylim = c(40, 90)) +` and setting the y limits to something you consider appropriate.

 

- Try some other [themes](https://ggplot2.tidyverse.org/reference/ggtheme.html). (at least try `theme_gray`, `theme_classic`, and `theme_minimal`).

 

Be prepared to submit your final RMarkdown file.

 

<!-- Based on: https://cs.calvin.edu/courses/data/202/21fa/ex/ex02/ex02-plotting-inst.html -->
