---
title: "Stat 245 -- HW Int Search Selection/Prediction"
author: "Adham Rishmawi"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(tidyverse)   # this loads mosaic, ggformula, etc. too
require(ggformula)
require(mosaic)
require(glmmTMB)
require(MuMIn)

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw(base_size=12))     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  echo = TRUE,      # for homework, always show R code (this is the default)
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small",   # slightly smaller font for code
  message = FALSE, warning = FALSE) # don't print warnings or messages in compiled document. So you MUST check them in RStudio!
```


```{r}
search1 <- read_csv('https://sldr.netlify.app/data/election_searches.csv')

```
```{r}
search_nb2 <- glmmTMB(Searches ~ Race_Ethnicity + Vote_Sway, 
                      data = search1,
                      family = nbinom2(link = 'log'))
summary(search_nb2)
```

```{r}
hyp_data1 <- expand.grid(Race_Ethnicity = "NA_NA_Hispanic_NA_NA_NA",
                         Vote_Sway = seq(from = -20, to = 20, by = 1)
                         )
hyp_data1
```
This potential dataset is obtaining data from the two constant variables i used. I used NA_NA_Hispanic_NA_NA_NA and Vote_Sway from -20 and 20.

###Predictions

```{r}
search_preds1 <- predict(search_nb2,
                         newdata = hyp_data1,
                         type = 'link',
                         se.fit = TRUE)
```


```{r}
hyp_data1 <- hyp_data1 |>
  mutate(pred = exp(search_preds1$fit),
         ci_low = exp(search_preds1$fit - 1.96*search_preds1$fit),
         ci_high = exp(search_preds1$fit + 1.96*search_preds1$fit))
```


```{r}
gf_line(pred ~ Vote_Sway,
        data = hyp_data1) |>
  gf_ribbon(ci_low + ci_high ~ Vote_Sway) 
  
```


 This prediction plot has completely failed us and shows us a plot with high uncertainty. I would conclude that this variable is probably not the most viable option. I would try something different from Race_Ethnicity.
 
 
###Model Selection

```{r}
search_nb2 <- search_nb2 |>
update(na.action = 'na.fail')
search1_dred <- dredge(search_nb2, rank = 'AIC')

search1_dred
```

I can conclude that my variable Race_ethnicity was the wrong choice because it failed my prediction plot. I should search for a different variable potentially like Sex or or Group as a better variable of choice. Our AIC isnt much affected from our variable so this even reinforces our decision to change or variables.  

 
 
 
 