---
title: "Stat 245 -- HW Internet Search Models"
author: "Adham Rishmawi"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 8
    fig_width: 7
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(tidyverse)   # this loads mosaic, ggformula, etc. too
require(ggformula)
require(mosaic)

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw(base_size=12))     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  echo = TRUE,      # for homework, always show R code (this is the default)
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small",   # slightly smaller font for code
  message = FALSE, warning = FALSE) # don't print warnings or messages in compiled document. So you MUST check them in RStudio!
```

But today, your task is to fit a regression model to understand how the number of internet searches someone does per week relates to demographic information. ->
I will do this by making my response variable Searches and my predictors Race _Ethnicity


```{r}
search <- read_csv('https://sldr.netlify.app/data/election_searches.csv')

```
```{r}
slr <- lm(Searches ~ Race_Ethnicity + Vote_Sway,
          data = search)
summary(slr)
```


By Looking at the r squared value we see that the predictors are not at all manipulating our response so we already have a negative sign




```{r}
s245::gf_acf(~slr)|>
gf_lims(y=c(-0.2,0.2))|>
gf_labs( title = "Independence Of Residuals", )
```


FAILED(two lines surpass the dashed areas meaning this test failed. Under lenient conditions, we could say this passed but for this I would still stick to failed!)

```{r}
gf_point(resid(slr)~fitted(slr))|>
gf_labs(x = 'Fitted Values', y = 'Residuals')|>
gf_labs(title = "Non-constant Variance Test", )
```



FAILED


```{r}
gf_point(log(Searches) ~ Race_Ethnicity,
data = search)|>
gf_labs(title = "Lack of non-linearity Test", )
```


I would say it is mostly non-linear, spread out recklessly though. I can not determine what type of slope this but i would say gracely, that this test passed.  


```{r}
search <- search |>
  mutate(preds = predict(slr),
         resids = resid(slr))

gf_histogram(~resids, data = search, bins = 35)|>
  gf_labs(title = "Histogram: Normality Of Residuals", x = "Residuals")


```


FAILED (there is a huge skew right which makes this histogram fail badly)

```{r}
library(glmmTMB)
search_nb2 <- glmmTMB(Searches ~ Race_Ethnicity + Vote_Sway, 
                      data = search,
                      family = nbinom2(link = 'log'))
```


```{r}
library(DHARMa)
nb2_sim <- simulateResiduals(search_nb2)
plotResiduals(nb2_sim,
              quantreg = FALSE)
```

This Graph looks problematic because the dashed line is jumping from above and below the 0.50 point which is something to take into consideration!


Overall i would say the Assessments failed and that we would need to start from the beginning because we need different predictors and response variable.
  