---
title: "dplyr Demo"
author: "Keith VanderLinden"
date: "Spring, 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document demonstrates the key `dplyr` functions in the context of a hotel-bookings dataset.

```{r packages, warning = FALSE, message = FALSE}
library(tidyverse)
```

## Dataset: Hotel Bookings

We&rsquo;d like to analyze hotel bookings data from [TidyTuesday Hotels](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-02-11/readme.md). Note that we&rsquo;ve converted the cancellation field into a factor, which will be useful in our analysis.

```{r}
hotels <- read_csv("data/hotels.csv", 
    col_types = cols(is_canceled = col_factor(levels = c("0", "1")))
    )
glimpse(hotels)
```

Notes:
- This dataset is too big to list easily. There are too many:
  - Rows (i.e., records, observations)
  - Columns (i.e., fields, variables)
- To do analysis, we need to "zoom in" on the data of interest.
- We read the `is_canceled` (stored as a double, values 1-cancelled; 0-not-cancelled) as a factor.

## `dplyr` Provides Data-Wrangling Functions

We now demonstrate some of the basic `dplyr` functions.

### Selecting Columns

The `select()` function allows us to zoom in on particular columns.

```{r}
hotels %>%
  select(lead_time, hotel)
```

Notes:
- You can add or remove column names.

### Filtering Rows

The `filter()` function allows us to zoom in on particular rows.

```{r}
hotels %>%
  filter(hotel == "Resort Hotel") 
```

Notes:
- You can add more filter conditions, e.g., `lead_time < 7`)
- The logical operators are:  `== < <= > >= !=`

### Summarizing Data

The `summarize()` function allows us to compute a variety of descriptive statistics, including counting, measurements of central tendency (e.g., mean, median, quartiles), etc.

```{r}
hotels %>%
  summarize(avg_lead_time = mean(lead_time))
```

Summarize is often used with grouping. Here we count the books in different market segments.

```{r}
hotels %>%
  group_by(market_segment) %>%
  summarize(booking_count = n())
```

### Arranging Rows

The `arrange()` function allows us to reorder the rows.

```{r}
hotels %>%
  group_by(market_segment) %>%
  summarize(average_lead_time = mean(lead_time)) %>%
  arrange(desc(average_lead_time))
```

Notes:
- Reverse the ordering using `desc()`.

### Mutating Structure

The `mutate()` function allows us to restructure the data.

```{r}
hotels %>%
  mutate(num_nights = stays_in_week_nights + stays_in_weekend_nights,
         week_nights = stays_in_week_nights,
         weekend_nights = stays_in_weekend_nights
         ) %>%
  arrange(desc(lead_time)) %>%
  select(num_nights,
         week_nights, 
         weekend_nights
         )  
```
Notes:
- We've focused on the data of interest by:
  - Computing a new field.
  - Renaming some fields.
  - Selecting the new/renamed fields.
- Be careful about ordering the pipelines, e.g., the sorting clause can't go just anywhere.

### Additional Functions

The `distinct()` function removes rows with duplicate values of the given columns.

```{r}
hotels %>%
  distinct(hotel) %>%
  arrange(hotel)
```
Note:
- You can count on lists of fields (e.g., add `market_segment`).

The `count()` function counts the rows.

```{r}
hotels %>%
  count(market_segment) 
```

Notes:
- `count(X)` is shorthand for `group_by(X) %>% summarize(n = n())`


# Using Functions in Combination

As we&rsquo;ve seen above, `dplyr` functions can be used in combination. 

The following `dplyr` pipeline plots counts that `geom_bar()` computes automatically.

```{r}
# manual pipeline
hotels %>%
  group_by(hotel) %>%
  summarize(count = n()) %>%
  ggplot() +
  aes(x = hotel, y = count) %>%
  geom_col()

# geom_bar only
hotels %>%
  ggplot() +
  aes(x = hotel) + 
  geom_bar()
```

The order of the wrangling functions in a script can matter. For example, this script fails:

```{r}
# Fails
# hotels %>%
#   select(hotel, children) %>%
#   filter(adults == 0)

# Works
hotels %>%
  filter(adults == 0) %>%
  select(hotel, children)
```

## Summary

`dplyr` provides a variety of data transformation functions that are useful in data wrangling.

See the RStudio [dplyr cheat sheet](https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-transformation.pdf).
