---
title: "DATA 202 Homework 4"
author: "Adham Rishmawi"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    code_download: yes
---

```{r setup, include=FALSE}
# Set some useful chunk options for code chunks.
knitr::opts_chunk$set(
  echo = TRUE,
  error = TRUE,    # display errors but keep going.
  comment = "",    # don't add '##' characters before output.
  message = FALSE  # don't include messages in the knitted output (check them in RStudio)
  )
library(tidyverse)
theme_set(theme_bw())
```


##importing data
```{r}
data1 <- read.csv("~/data202/homework4/country-population.csv")
data2 <- read.csv("~/data202/homework4/country-codes.csv")

```



##merging datasets
```{r}
joined_data <- merge(data1, data2, by.x = "country", 
              all.x = TRUE, all.y = FALSE)

```


####filtering bad instances

```{r}
df2 <- joined_data[,!names(joined_data) %in% c("code")]

```







##Query Raw Population Data
```{r}
staff4 <-df2 |>
  group_by(`country`,`X1800`,`X1880`)

```
 
 
##Exploring the data

Initial question: do populations in European countries increase at a less rate then Asian countries


## plot

We added a column by mutating a new one in and calling it change and making it the difference between the last recorded population and the first recorded
```{r}
staff4 <-df2 |>
  group_by(`country`,`X1800`,`X1880`) |>
  mutate(change = `X2015`-`X1800`) |>
  mutate(chnage_upper = `X2015` - `X2010`) |>
  mutate(change_lower = `X1850` - `X1800`)

staff4
```
We Can see that the population increase for a majority of Asian countries has sky rocketed over years and in both sub sections of change_lower and change_higher. This could be due to the fact that Asian countries typically have a larger family size then compared to the ones found in European countries. The change_higher is much smaller in European countries then it is in Asian countries which indicates to the observer of this plot that Asian countries have maintained a higher population in recent years and almost over every year. 



## Conclusion
My initial question/statement aligns with the truth and I am able to conclude that Asian countries are increasing faster in population then European countries!
