---
title: "Stat 245 -- HW ??"
author: "A. Student"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    fig_height: 4
    fig_width: 8
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(tidyverse)   # this loads mosaic, ggformula, etc. too
require(ggformula)
require(mosaic)

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw(base_size=12))     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  echo = TRUE,      # for homework, always show R code (this is the default)
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small",   # slightly smaller font for code
  message = FALSE, warning = FALSE) # don't print warnings or messages in compiled document. So you MUST check them in RStudio!
```


```{r}
daily_rides <- read_csv("~/data202/homework02/bikeshare-day.csv") %>%
    mutate(across(c(season, year, holiday, workingday, day_of_week, weather_type, rider_type), as.factor))
    glimpse(daily_rides)
```
```{r}
daily_rides <- daily_rides %>%
  mutate(day_of_week = factor(day_of_week, levels = c(0, 1, 2, 3, 4, 5, 6), labels = c("Sun", "Mon", "Tues", "Weds", "Thurs", "Fri", "Sat")))
```
##### By looking at the days of the weekend and weekdays we can determine that 0 is Sunday and the following days must be the chronological order

date: this stamps the marker of when the data was recorded/

rider_type: declares what type of rider this person is and how to differentiate between casual rider and heavy user.

rides: the amount of times a ride has occurred for this particular individual.

holiday: is a binary option to determine whether rides occurred on a holiday or not.

working day: is also a two option criteria where it determines whether  the day is a weekend or weekday.

day_of_week: determines which day it is from Sunday to Saturday.


```{r}
daily_rides %>%
  ggplot() +
  aes(x = date, y = rides, col = rider_type) +
  geom_point(size=1, alpha= 0.5) +
  geom_smooth() +
  labs(
    x = "Date", 
    y = "number of rides", 
    title = "Rides by date",
    subtitle = "for casual and registered riders")


```
```{r}
daily_rides %>% 
  ggplot() +
  aes(x = temp, y = rides, color = rider_type) +
  facet_wrap(~workingday) +
  geom_point()
```
###### this graph is split between weekend and workdays to demonstrate the neccesity of riding a bike and see the discrepancy  
```{r}
daily_rides %>% 
  ggplot() +
  aes(x = temp, y = rides, color = rider_type) +
  facet_wrap(~rider_type) +
  geom_point()
```
##### this graph is facet of casual vs registered indivuals to tell the observer that registered bike user willuse their bikes in lower temperatures


```{r}
daily_rides %>% 
  ggplot() +
  aes(x = day_of_week, y = rides, fill = rider_type) +
  labs(
    x = "Day of Week", 
    y = "Rides", 
    title = "Graph#5",
    subtitle = "best attempt") + 
   geom_boxplot()
theme_classic()
```
This plot shows us the big grouping of people and whether that be registered or casual people and their amount of rides!


```{r}
daily_rides %>% 
  ggplot() +
  aes(x = rides, y = humidity, color =   rider_type) +
  labs(
    x = "Rides", 
    y = "Humidity level", 
    title = "Graph#5",
    subtitle = "best attempt") + 
   geom_col()
```

this graph shows us the that registered riders tend to ride in lower humidified weathers then casual riders! 


  